A comparison of design-based and model-based approaches for spatial data.

### 2021-09-17 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.Rmd
added simulation parameter tables and graph of lognormal distribution.

There were several variables to alter in the simulations, and the names of the scenarios in future plots mirror these variables

* correlation type: dependent errors or independent errors
* error type:
    * normal: mean 0, variance 2
    * lognormal: log scale mean 0, log scale variance 2 (total variance 47)
    * lognormalbig: log scale mean 0, log scale variance 4 (total variance 2,926)
* sample sizes: $n = 10, 50, 150; N = 900$
* layout: gridded vs random uniform population locations confined to a 1 x 1 unit square

>>>>

We used a numerical simulation study to investigate performance of four design-analysis combinations, summarized in Table \ref{tab:designanalysis}.

```{r, echo = FALSE}
tab <- matrix(c("IRS-Design", "IRS-Model", "GRTS-Design", "GRTS-Model"),
       nrow = 2) %>%
  as.table()
row.names(tab) <- c("Design", "Model")
library(knitr)
kable(tab, col.names = c("IRS", "GRTS"), caption = "\\label{tab:designanalysis} Types of Sampling Design and Analysis combinations considered in the simulation study. The columns give the two types of sampling designs while the rows give the two types of analyses.")
## I really want "Analysis" to appear as a row header and "Sampling Design" 
## to appear as a column header, but this isn't easy to figure out.
## I couldn't figure this out with xtable either so it might need to be
## made "by hand."
```

We used a crossed design with the simulation parameters given in Table \ref{tab:parmtab} for a total of 36 scenarios. All scenarios used exponential correlation with an effective range of $\sqrt{2}$ for $N = 900$ response values simulated on the unit square in either random locations (Site Locations = Random) or gridded locations (Site Locations = Gridded). For the lognormal scenarios, the response values were simulated using the specified correlation parameters using a normal distribution and were subsequently exponentiated. A total variance of 2 and a mean of 0 on the normal scale is equivalent to a total variance of 47 and a mean of 2.72 after exponentiation. Therefore, when the model-based methods were used, the correlation is mis-specified. We chose to simulate values with a lognormal distribution so that we could test the model-based analysis approach with a mis-specified model and so that we could test both analysis approaches on data that exhibits a large amount of skewness (Figure \ref{fig:figlog}).

```{r}
n <- c("50", "100", "200")
site_locations <- c("Random", "Gridded", "")
psill_ratio <- c("0", "0.5", "0.9")
resp_type <- c("Normal", "Lognormal", "")
parm_tab <- rbind(n, site_locations, psill_ratio, resp_type)
row.names(parm_tab) <- c("Sample Size (n)", "Site Locations", 
                         "Partial Sill Variance Ratio", "Response Type")
kable(parm_tab, col.names = c("", "", ""),
      caption = "\\label{tab:parmtab} Simulation parameters. Total variability for all scenarios was 2 so that the partial sill was 0, 1, or 1.8.")
```

```{r, figlog, out.width = "100%", fig.cap = "Lognormal distribution with a mean of 0 on the log scale and variance of 2 on the log scale. The distribution exhibits moderate to strong right-skewness."}
mu <- 0; var <- 2
xvar <- seq(0.001, 10, by = 0.01)
yvar <- dlnorm(xvar, meanlog = 0, sdlog = sqrt(var))
df <- tibble(xvar = xvar, yvar = yvar)
ggplot(df, aes(x = xvar, y = yvar)) +
  geom_line()
```

* correlation type: dependent errors or independent errors
* error type:
    * normal: mean 0, variance 2
    * lognormal: log scale mean 0, log scale variance 2 (total variance 47)
    <!-- * lognormalbig: log scale mean 0, log scale variance 4 (total variance 2,926) -->
* sample sizes: $n = 10, 50, 150; N = 900$
* layout: gridded vs random uniform population locations confined to a 1 x 1 unit square

### 2021-09-17 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
added simulation parameter tables and graph of lognormal distribution.

There were several variables to alter in the simulations, and the names
of the scenarios in future plots mirror these variables

@@

  \begin{itemize}
  \tightlist
  \item
    normal: mean 0, variance 2
  \item
    lognormal: log scale mean 0, log scale variance 2 (total variance
    47)
  \item
    lognormalbig: log scale mean 0, log scale variance 4 (total variance
    2,926)
  \end{itemize}
\item
  sample sizes: \(n = 10, 50, 150; N = 900\)
\item
  layout: gridded vs random uniform population locations confined to a 1
  x 1 unit square
\end{itemize}

@@

\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-3-1.pdf}

@@

\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-4-1.pdf}

@@

>>>>

We used a numerical simulation study to investigate performance of four
design-analysis combinations, summarized in Table
\ref{tab:designanalysis}.

\begin{longtable}[]{@{}lll@{}}
\caption{\label{tab:designanalysis} Types of Sampling Design and
Analysis combinations considered in the simulation study. The columns
give the two types of sampling designs while the rows give the two types
of analyses.}\tabularnewline
\toprule
& IRS & GRTS \\
\midrule
\endfirsthead
\toprule
& IRS & GRTS \\
\midrule
\endhead
Design & IRS-Design & GRTS-Design \\
Model & IRS-Model & GRTS-Model \\
\bottomrule
\end{longtable}

We used a crossed design with the simulation parameters given in Table
\ref{tab:parmtab} for a total of 36 scenarios. All scenarios used
exponential correlation with an effective range of \(\sqrt{2}\) for
\(N = 900\) response values simulated on the unit square in either
random locations (Site Locations = Random) or gridded locations (Site
Locations = Gridded). For the lognormal scenarios, the response values
were simulated using the specified correlation parameters using a normal
distribution and were subsequently exponentiated. A total variance of 2
and a mean of 0 on the normal scale is equivalent to a total variance of
47 and a mean of 2.72 after exponentiation. Therefore, when the
model-based methods were used, the correlation is mis-specified. We
chose to simulate values with a lognormal distribution so that we could
test the model-based analysis approach with a mis-specified model and so
that we could test both analysis approaches on data that exhibits a
large amount of skewness (Figure \ref{fig:figlog}).

\begin{longtable}[]{@{}llll@{}}
\caption{\label{tab:parmtab} Simulation parameters. Total variability
for all scenarios was 2 so that the partial sill was 0, 1, or
1.8.}\tabularnewline
\toprule
& & & \\
\midrule
\endfirsthead
\toprule
& & & \\
\midrule
\endhead
Sample Size (n) & 50 & 100 & 200 \\
Site Locations & Random & Gridded & \\
Partial Sill Variance Ratio & 0 & 0.5 & 0.9 \\
Response Type & Normal & Lognormal & \\
\bottomrule
\end{longtable}

\begin{figure}
\includegraphics[width=1\linewidth]{SpatialDVM_Manuscript_files/figure-latex/figlog-1} \caption{Lognormal distribution with a mean of 0 on the log scale and variance of 2 on the log scale. The distribution exhibits moderate to strong right-skewness.}\label{fig:figlog}
\end{figure}

@@

  \begin{itemize}
  \tightlist
  \item
    normal: mean 0, variance 2
  \item
    lognormal: log scale mean 0, log scale variance 2 (total variance
    47)
  \end{itemize}
\item
  sample sizes: \(n = 10, 50, 150; N = 900\)
\item
  layout: gridded vs random uniform population locations confined to a 1
  x 1 unit square
\end{itemize}

@@

@@

@@

\includegraphics{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-8-1.pdf}
